---
output: github_document
---

```{r, include = FALSE}
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")
set.seed(1)
options(datatable.print.class = FALSE, datatable.print.keys = FALSE, width = 200)
```

# mlr3tuning

Package website: [release](https://mlr3tuning.mlr-org.com/) | [dev](https://mlr3tuning.mlr-org.com/dev/)

<!-- badges: start -->
[![tic](https://github.com/mlr-org/mlr3tuning/workflows/tic/badge.svg?branch=main)](https://github.com/mlr-org/mlr3tuning/actions)
[![CRAN Status](https://www.r-pkg.org/badges/version-ago/mlr3tuning)](https://cran.r-project.org/package=mlr3tuning)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
[![CodeFactor](https://www.codefactor.io/repository/github/mlr-org/mlr3tuning/badge)](https://www.codefactor.io/repository/github/mlr-org/mlr3tuning)
<!-- badges: end -->

This package provides hyperparameter tuning for [mlr3](https://mlr3.mlr-org.com).
It offers various tuning methods e.g. grid search, random search and generalized simulated annealing and iterated racing and different termination criteria can be set and combined.
`AutoTuner` provides a convenient way to perform nested resampling in combination with `mlr3`.
The package is build on [bbotk](https://github.com/mlr-org/bbotk) which provides a common framework for optimization.

## Resources

- mlr3book chapters on [tuning search spaces](https://mlr3book.mlr-org.com/optimization.html#searchspace), [hyperparameter tuning](https://mlr3book.mlr-org.com/optimization.html#tuning) and [nested resampling](https://mlr3book.mlr-org.com/optimization.html#nested-resampling).
- mlr3gallery [posts](https://mlr3gallery.mlr-org.com/#category:mlr3tuning)
- [cheatsheet](https://cheatsheets.mlr-org.com/mlr3tuning.pdf)

## Installation

Install the last release from CRAN:

```{r eval = FALSE}
install.packages("mlr3tuning")
```

Install the development version from GitHub:

```{r eval = FALSE}
remotes::install_github("mlr-org/mlr3tuning")
```

## Example

```{r, include = FALSE}
# mute load messages
library("mlr3tuning")
```

### Basic hyperparameter tuning

```{r}
library("mlr3tuning")

# retrieve task
task = tsk("pima")

# load learner and set search space
learner = lrn("classif.rpart", cp = to_tune(1e-04, 1e-1, logscale = TRUE))

# hyperparameter tuning on the pima indians diabetes data set
instance = tune(
  method = "random_search",
  task = task,
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.ce"),
  term_evals = 10,
  batch_size = 5
)

# best performing hyperparameter configuration
instance$result

# all evaluated hyperparameter configuration
as.data.table(instance$archive)

# fit final model on complete data set
learner$param_set$values = instance$result_learner_param_vals
learner$train(task)
```

### Automatic tuning

```{r}
# retrieve task
task = tsk("pima")

# construct auto tuner
at = auto_tuner(
  method = "random_search",
  learner = lrn("classif.rpart", cp = to_tune(1e-04, 1e-1, logscale = TRUE)),
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.ce"),
  term_evals = 10,
  batch_size = 5
)

# train/test split
train_set = sample(task$nrow, 0.8 * task$nrow)
test_set = setdiff(seq_len(task$nrow), train_set)

# tune hyperparameters and fit final model on the complete data set in one go
at$train(task, row_ids = train_set)

# best performing hyperparameter configuration
at$tuning_result

# all evaluated hyperparameter configuration
as.data.table(at$archive)

# predict new data with model trained on the complete data set and optimized hyperparameters
at$predict(task, row_ids = test_set)
```

### Nested resampling

```{r}
# retrieve task
task = tsk("pima")

# load learner and set search space
learner = lrn("classif.rpart", cp = to_tune(1e-04, 1e-1, logscale = TRUE))

# nested resampling
rr = tune_nested(
  method = "random_search",
  task =  task,
  learner = learner,
  inner_resampling = rsmp("holdout"),
  outer_resampling = rsmp("cv", folds = 3),
  measure = msr("classif.ce"),
  term_evals = 10,
  batch_size = 5
)

# aggregated performance of all outer resampling iterations
rr$aggregate()

# performance scores of the outer resampling
rr$score()

# inner resampling results
extract_inner_tuning_results(rr)

# inner resampling archives
extract_inner_tuning_archives(rr)
```
