---
output: github_document
---

```{r, include = FALSE}
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")
set.seed(1)
options(datatable.print.class = FALSE, datatable.print.keys = FALSE)
```

# mlr3tuning

Package website: [release](https://mlr3tuning.mlr-org.com/) | [dev](https://mlr3tuning.mlr-org.com/dev/)

<!-- badges: start -->
[![tic](https://github.com/mlr-org/mlr3tuning/workflows/tic/badge.svg?branch=main)](https://github.com/mlr-org/mlr3tuning/actions)
[![CRAN Status](https://www.r-pkg.org/badges/version-ago/mlr3tuning)](https://cran.r-project.org/package=mlr3tuning)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
[![CodeFactor](https://www.codefactor.io/repository/github/mlr-org/mlr3tuning/badge)](https://www.codefactor.io/repository/github/mlr-org/mlr3tuning)
<!-- badges: end -->

This package provides hyperparameter tuning for [mlr3](https://mlr3.mlr-org.com).
It offers various tuning methods e.g. grid search, random search and generalized simulated annealing and different termination criteria can be set and combined.
`AutoTuner` provides a convenient way to perform nested resampling in combination with `mlr3`.
The package is build on [bbotk](https://github.com/mlr-org/bbotk) which provides a common framework for optimization.

## Resources

- mlr3book chapters on [tuning search spaces](https://mlr3book.mlr-org.com/searchspace.html), [hyperparameter tuning](https://mlr3book.mlr-org.com/tuning.html) and [nested resampling](https://mlr3book.mlr-org.com/nested-resampling.html).
- mlr3gallery [posts](https://mlr3gallery.mlr-org.com/#category:mlr3tuning)
- [cheatsheet](https://cheatsheets.mlr-org.com/mlr3tuning.pdf)

## Installation

Install the last release from CRAN:

```{r eval = FALSE}
install.packages("mlr3tuning")
```

Install the development version from GitHub:

```{r eval = FALSE}
remotes::install_github("mlr-org/mlr3tuning")
```

## Example

```{r}
library("mlr3")
library("mlr3tuning")
library("paradox")

task = tsk("pima")
learner = lrn("classif.rpart")
resampling = rsmp("holdout")
measure = msr("classif.ce")

# Create the search space with lower and upper bounds
learner$param_set$values$cp = to_tune(0.001, 0.1)
learner$param_set$values$minsplit = to_tune(1, 10) 

# Define termination criterion
terminator = trm("evals", n_evals = 20)

# Create tuning instance
instance = TuningInstanceSingleCrit$new(
  task = task,
  learner = learner,
  resampling = resampling,
  measure = measure,
  terminator = terminator)

# Load tuner
tuner = tnr("grid_search", resolution = 5)

# Trigger optimization
tuner$optimize(instance)
```
