---
title: "Introduction to Fitness Functions"
author: "Daniel Schalk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Fitness Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(mlr3tuning)
knitr::opts_knit$set(
  datatable.print.keys = FALSE,
  datatable.print.class = TRUE
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(123)
```

`ml3tuning` is an extension of `mlr3` that includes tuning.

## Define Required Objects 



```{r}
mlr_tasks
# list keys
mlr_tasks$keys()
# get a quick overview
as.data.frame(mlr_tasks)
```

Here, we will stick to the popular [iris data set](https://en.wikipedia.org/wiki/Iris_flower_data_set):

```{r}
task = mlr_tasks$get("iris")
print(task)
```

Several task properties and characteristics can be queried using the task's methods (see `mlr3::Task`).
Most of them should be self-explanatory.

```{r}
task$nrow
task$ncol
task$target_names
task$class_n
task$class_names
task$feature_names
task$formula
task$head(n = 3)
```

In `mlr3`, each row has a unique identifier (row name) which can be either `integer` or `character.`
These can be used to select specific rows.

```{r}
# iris uses integer row_ids
task$row_ids
# filter rows 1, 51, and 101
task$data(rows = c(1, 51, 101))
# filter rows 1, 51, and 101 and only select column "Species"
task$data(rows = c(1, 51, 101), cols = "Species")
```
